Climate indicators

### 2023-02-27 DRAFT_climate_indicators.Rmd
ready to test real data

# Cilmate indicators

@@

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
library(microbenchmark)
dir <- substr(getwd(), 1,2)

@@

This chaters describes a workflow for generating or preparing indicators based on interpolated climate data from [SeNorge](https://senorge.no/).

@@

The data includes the last normal period (1961-1990) which defines the reference condition for climate variables. Therefore the temporal resolution is very good. Also considering tha daily resolution of the data.

@@

Varied

@@

```{r, warning=F}
tg_1957 <- stars::read_ncdf(paste0(path, "/rr_tm/seNorge2018_1957.nc"))
```

@@

... and four attributes

@@

Precipitation (rr), mean temperature (tg), max temperatur (tx) and min temperature (tn) are included in the same file.

For this example workflow I will only focus on tg (temperature). I could subset the data like this

```{r}
tg_1957 %>% select(tg)
```

, but since this is a proxy object, it will not initiate any change untill later, when I write to file and all the lazy operations are done in squeal. Therefore, I will import a test data set, which is smaller and which I can import to memory. Then I can perform the operations on that data set and we can see the results.

@@

```{r}
dummy_aa <- x %>%
  filter(time == oldTime[1]) %>%
  adrop() %>%
  mutate(accuntingAreas = rep(c(1,2), each=length(Attribute_A)/2)) %>%
  select(accuntingAreas)

@@

```{r}
dummy_et <- x %>%
  filter(time == oldTime[1]) %>%
  adrop() %>%
  mutate(ecosystemType = rep(c(1,NA), length.out = length(Attribute_A))) %>%
  select(ecosystemType)

@@

I need to code a for loop for importing the .nc files.

For then selecting the attribute, simply do like this:

```{r}
# list attribute names
names(x)
```

```{r}
x_sub <- x %>% 
  select(Attribute_A)
names(x_sub)
```

@@

```{r, fig.cap="Demonstrating the effecty of maskig the dummy data using a perfectly aligne raster ET map."}
x_masked <- x_aug
x_masked[is.na(dummy_et)] <- NA
plot(x_masked[,,,1])
```

```{r withMasking, fig.cap="Compareing computation time before and after masking of raster data. There is no performance increase by masking before aggregating."}
autoplot(
  microbenchmark(
No_masking =  x_aug %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names)),
Masking =  x_masked %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names))
 ), times=30
)

@@

Since there is no increase in performance from masking, as we see in Fig. \@ref(fig:withMasking) then it basically means we it is slower to mask beforehand, since the masking was not past of the benchmark assessment.

@@

>>>>

# Climate indicators

@@

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
library(microbenchmark)
library(ncmeta)
dir <- substr(getwd(), 1,2)

@@

This chapters describes a workflow for generating or preparing indicators based on interpolated climate data from [SeNorge](https://senorge.no/).

@@

The data includes the last normal period (1961-1990) which defines the reference condition for climate variables. Therefore the temporal resolution is very good. Also considering the daily resolution of the data.

@@

Varied. Specified below for each parameter.

@@

```{r, warning=F}
tg_1957 <- stars::read_ncdf(paste0(path, "/rr_tm/seNorge2018_1957.nc"),
                            var="tg")
```

@@

Initially the data had four attributes, but I subsettet to only include _tg_.

@@

`tg_1957` is a stars proxy object and [most commands](https://r-spatial.github.io/stars/articles/stars8.html) will not initiate any change untill later, typically I write to file and all the lazy operations are done in squeal. Therefore, I will prepare a test data set, which is smaller and which I can import to memory. Then I can perform the operations on that data set and we can see the results.

@@

```{r, fig.cap="Showing the example account area delineation (raster format)"}
dummy_aa <- x %>%
  filter(time == oldTime[1]) %>%
  adrop() %>%
  mutate(accuntingAreas = rep(c(1,2), each=length(Attribute_A)/2)) %>%
  select(accuntingAreas)

@@

```{r, fig.cap="Showing the example Ecosystem type delineation data."}
dummy_et <- x %>%
  filter(time == oldTime[1]) %>%
  adrop() %>%
  mutate(ecosystemType = rep(c(1,NA), length.out = length(Attribute_A))) %>%
  select(ecosystemType)

@@

I include a for loop for importing the .nc files at the end.

@@

```{r, fig.cap="Demonstrating the effect of maskig the dummy data using a perfectly aligne raster ET map."}
x_masked <- x_aug
x_masked[is.na(dummy_et)] <- NA
plot(x_masked[,,,1])
```

```{r withMasking, fig.cap="Comparing computation time before and after masking of raster data. There is no performance increase by masking before aggregating."}
autoplot(
  microbenchmark(
No_masking =  x_aug %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names)),
Masking =  x_masked %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names))
 ), times=30
)

@@

Since there is no increase in performance from masking, as we see in Fig. \@ref(fig: withMasking) then it basically means we it is slower to mask beforehand, since the masking was not past of the benchmark assessment.

@@

Now let's try it on the real data

```{r}
path <- path <- ifelse(dir == "C:", 
      "R:/",
      "/data/R/")

path2 <- paste0(path, "GeoSpatialData/Meteorology/Norway_SeNorge2018_v22.09/Original/rr_tm/") #seNorge2018_1957.nc

myFiles <- list.files(path2, full.names = T)[1:2]
start_month <-  "06"
end_month <- "08"
real_temp_summer <- NULL

for(i in 1:length(myFiles)){

  temp <- stars::read_ncdf(paste(path2, var="tg"))
  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  start <- ym(paste(year_temp, start_month, sep="-"))
  end <-  ym(paste(year_temp, end_month, sep="-"))
  myInterval <- interval(start, end)

  summerTemp_fullSeries <- c(temp, summerTemp_fullSeries)
  rm(temp)
}
```
